Measurement and Probability in Fuzzy Quantum Theories

### 2017-02-01 prop.tex
prop.tex \section{Quantum and Classical} Finish convex question...

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
an interval of probability $[l_{2},r_{2}]$, what is the interval
of probability for the event $E_{1}\cup E_{2}$? The answer is somewhat
subtle: although it is possible to use the sum of the intervals $[l_{1}+l_{2},r_{1}+r_{2}]$
as the combined probability, one can find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is evidence supporting the complement of $E$. The
complement of $E$ must therefore have probability $\left[1-r,1-l\right]$
which we abbreviate $\left[1,1\right]-\left[l,r\right]$. Given a
sample space~$\Omega$, its set of events~$\events$, and a collection
of intervals $\mathscr{I}$, a function~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is a classical interval-valued probability measure if and only if
$\bar{\mu}$ satisfies the following conditions~\cite{JamisonLodwick2004}
where the last line uses $\subseteq$ to allow for tighter intervals
that exploit the complement event: 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$. 
\item $\bar{\mu}(\Omega)=[1,1]$. 
\item For any event $E$, $\bar{\mu}\left(\Omega\backslash E\right)=\left[1,1\right]-\bar{\mu}\left(E\right)$ 
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have $\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)$. 
\end{itemize}

@@

Looking closely at the example above, we might think that the
probability measure $\bar{\mu}$ is induced by the state $\ket{2}$
because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$, or we
might think it is induced by the state $\ket{1}$ because
$\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$, or yet we might
think it is induced by the mixed state $\frac{\mathbb{1}}{3}$ because
$\bar{\mu}(\proj{\phi})\ne\necess$ for all $\ket{\phi}$. Each of these
possibilities is compatible with some but not all of the
observations. All is not lost however: if the observations are made
more precise, we conjecture that some of the inconsistencies
disappear. A partial proof of this conjecture is the following example
which refines the previous probability measure by adding more precise
intervals. It remains to prove that this refined measure is convex,
however.

@@

\section{Subset of Commuting Events}

\footnote{\yutsung{The Kochen-Specker theorem is about assigning definitely
value to observables such that the values of commuting observables
are consistent; the Bell theorem is about entangled states in different
locations. If our results are more related to commuting observables,
it is more related to the Kochen-Specker theorem than the Bell theorem...}}

Given an orthonormal basis~$\mathcal{B}=\left\{ \ket{\psi_{1}},\ldots,\ket{\psi_{d}}\right\} $
of a Hilbert space, the set of projections generated by the basis
is $\events'=\set{\sum_{\ket{\psi}\in S}\ket{\psi}}{S\subseteq\mathcal{B}}$.
Because $\left(2^{\mathcal{B}},\emptyset,\mathcal{B},\cup,\cap\right)$
is isomorphic to $\left(\events',\mathbb{0},\mathbb{1},\vee,\cdot\right)$
as a Boolean algebra, a quantum IVPM~$\bar{\mu}':\events'\rightarrow\mathscr{I}$
is essentially a classical IVPM~$\bar{\mu}:2^{\mathcal{B}}\rightarrow\mathscr{I}$.
By the Shapley theorem, if $\bar{\mu}'$ is convex, it should have
a non-empty core.

The problem is: if $\events'$ is only a commuting subset of $\events$,
does a convex quantum IVPM~$\bar{\mu}':\events'\rightarrow\mathscr{I}$
always has a non-empty core?

>>>>

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
an interval of probability $[l_{2},r_{2}]$, what is the interval
of probability for the event $E_{1}\cup E_{2}$? The answer is somewhat
subtle: although it is possible to use the sum of the intervals $[l_{1}+l_{2},r_{1}+r_{2}]$
as the combined probability, one can find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is evidence supporting the complement of $E$. The
complement of $E$ must therefore have probability $\left[1-r,1-l\right]$
which we abbreviate $\left[1,1\right]-\left[l,r\right]$. Given a
sample space~$\Omega$, its set of events~$\events$, and a collection
of intervals $\mathscr{I}$, a function~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is a classical interval-valued probability measure if and only if
$\bar{\mu}$ satisfies the following conditions~\cite{JamisonLodwick2004}
where the last line uses $\subseteq$ to allow for tighter intervals
that exploit the complement event: 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$. 
\item $\bar{\mu}(\Omega)=[1,1]$. 
\item For any event $E$, $\bar{\mu}\left(\Omega\backslash E\right)=\left[1,1\right]-\bar{\mu}\left(E\right)$ 
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have 
\begin{equation}
\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)\textrm{ .}\label{eq:classical-include}
\end{equation}
\end{itemize}

@@

Looking closely at the example above, we might think that the probability
measure $\bar{\mu}$ is induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or we might think it is induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$,
or yet we might think it is induced by the mixed state $\frac{\mathbb{1}}{3}$
because $\bar{\mu}(\proj{\phi})\ne\necess$ for all $\ket{\phi}$.
Each of these possibilities is compatible with some but not all of
the observations. All is not lost however: if the observations are
made more precise, we conjecture that some of the inconsistencies
disappear. A partial proof of this conjecture is the following example
which refines the previous convex probability measure by adding more
precise intervals.

@@

\section{Quantum and Classical}

\begin{lemma}Given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$
for a given Hilbert space~$\Hilb$ of dimension~$d$, and a quantum
IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$. Then, the function
$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ defined by
$\bar{\mu}^{\Omega}\left(E\right)=\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)$
is a classical IVPM, where $\sum_{\ket{j}\in\emptyset}\proj{j}$ is
defined to be $\mathbb{0}$.\end{lemma}
\begin{proof}
~
\begin{itemize}
\item $\bar{\mu}^{\Omega}(\emptyset)=\bar{\mu}\left(\sum_{\ket{j}\in\emptyset}\proj{j}\right)=\bar{\mu}\left(\mathbb{0}\right)=[0,0]$. 
\item $\bar{\mu}^{\Omega}(\Omega)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)=[1,1]$. 
\item For any event $E$, $\bar{\mu}^{\Omega}\left(\Omega\backslash E\right)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega\backslash E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}-\sum_{\ket{j}\in E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)-\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)=\left[1,1\right]-\bar{\mu}^{\Omega}\left(E\right)$. 
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have $\bar{\mu}^{\Omega}\left(\bigcup_{i=1}^{N}E_{i}\right)=\bar{\mu}\left(\sum_{\ket{j}\in\bigcup_{i=1}^{N}E_{i}}\proj{j}\right)=\bar{\mu}\left(\sum_{i=1}^{N}\sum_{\ket{j}\in E_{i}}\proj{j}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(\sum_{\ket{j}\in E_{i}}\proj{j}\right)=\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)$. 
\end{itemize}
\end{proof}
\footnote{\yutsung{The Kochen-Specker theorem is about assigning definitely
value to observables such that the values of commuting observables
are consistent; the Bell theorem is about entangled states in different
locations. If our results are more related to commuting observables,
it is more related to the Kochen-Specker theorem than the Bell theorem...}}\begin{lemma}Given a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
if $\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ is a classical
convex IVPM for any orthonormal basis~$\Omega$, then $\bar{\mu}$
is a quantum convex IVPM.\end{lemma}
\begin{proof}
Pick two commuting events~$P_{0},P_{1}\in\events$. Since they are
commuting, they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
Since they are projectors, the eigenvalues of them are $0$ or $1$
so that we can express them as $P_{0}=\sum_{\ket{j}\in E_{0}}\proj{j}$
and $P_{1}=\sum_{\ket{j}\in E_{1}}\proj{j}$. Then, 
\begin{eqnarray*}
 &  & \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\\
 & = & \bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\vee\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)+\bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)\\
 & = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cup E_{1}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cap E_{1}}\proj{j}\right)\\
 & = & \bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\\
 & \subseteq & \bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\\
 & = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\\
 & = & \bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)
\end{eqnarray*}
\end{proof}
\begin{lemma}When the number of elements in $\Omega$ is less than
or equal to 3, every classical convex IVPM~$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$.\end{lemma}
\begin{proof}
Given two events~$E_{0},E_{1}\subseteq\Omega$, we want to verify
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}\label{eq:classical-convex}
\end{equation}
When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). When
$E_{0}\cup E_{1}=\Omega$, we have $E_{0}^{c}\cap E_{1}^{c}=\emptyset$,
where $E_{0}^{c}=\Omega\backslash E_{0}$ and $E_{1}^{c}=\Omega\backslash E_{1}$.
Thus, 
\begin{eqnarray*}
 &  & \bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 & \Rightarrow & \left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 & \Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cup E_{1}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cap E_{1}^{c}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{1}^{c}\right)\right)\\
 & \Rightarrow & \bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}
\end{eqnarray*}
When the number of elements in $\Omega$ is less than or equal to
3, either $E_{0}\cap E_{1}=\emptyset$ or $E_{0}\cup E_{1}=\Omega$
must happen so that $\bar{\mu}^{\Omega}$is convex.
\end{proof}
\begin{thm}Given a Hilbert space~$\Hilb$ of dimension~$3$, every
quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is convex.\end{thm}
\begin{proof}
The direct consequence of the previous lemmas.
\end{proof}
